### 2021-01-15 ch03_esmvaltool/esmvaltool.tex
Added text and figures to section 4.1

To ensure a rapid evaluation of the \acs{CMIP}6 models, the \ac{ESMValTool} is
fully integrated into the infrastructure of the \ac{ESGF}, which provides the
\ac{CMIP} model output for the general public \autocite{Eyring2016b}. As soon
as new model data is published on the \ac{ESGF} servers, it can be accessed
with the \ac{ESMValTool} and analyzed. This instantaneous evaluation of the
\ac{CMIP} models is urgently needed since there is a growing dependency on
\ac{CMIP} products by a broad research community and by national and
international climate assessments. For this reason, the \ac{ESMValTool} is for
example used to evaluate climate model output and observational data in several
chapters of the upcoming \ac{AR6} of the \ac{IPCC}.

@@

\begin{table}[t]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Description \\
    \midrule
    Preprocessor & \emph{amplitude} & Amplitude of periodic phenomena (\eg{}
    cycles) \\
    \cmidrule(l){2-3}
    & \emph{land/sea fraction weighting} & Weighting of fields based on
    the land or sea fraction of the respective grid cells \\
    \cmidrule(l){2-3}
    & \emph{trend} & Linear trend (slope of linear regression) \\
    \midrule
    Derived variable & \emph{asr} & Absorbed shortwave radiation \\
    \cmidrule(l){2-3}
    & \emph{co2s} & Atmospheric \ac{CO2} concentration at
    surface \\
    \cmidrule(l){2-3}
    & \emph{et} & Evapotranspiration \\
    \cmidrule(l){2-3}
    & \emph{rlntcs} & Clear-sky net \acf{TOA} longwave radiation \\
    \cmidrule(l){2-3}
    & \emph{rsntcs} & Clear-sky net \ac{TOA} shortwave radiation \\
    \cmidrule(l){2-3}
    & \emph{uajet} & Position of austral jet stream \\
    \bottomrule
  \end{tabulary}
  \caption{Summary of new preprocessor functions and variable derivation
    scripts contributed to the \acf{ESMValCore} by the author of this thesis.}
  \label{tab:03:changes_to_esmvalcore}
\end{table}

@@

\begin{table}[t]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Main reference \\
    \midrule
    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
    & \emph{cox18nature} & \autocite{Cox2018} \\
    & \emph{ecs} & \autocite{Gregory2004} \\
    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
    & \emph{tcr} & \autocite{Gregory2008} \\
    \midrule
    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
    & CT2019 & \autocite{Jacobson2020} \\
    & GCP & \autocite{Friedlingstein2019} \\
    & HWSD & \autocite{Wieder2014} \\
    & JMA-TRANSCOM & \autocite{Maki2010} \\
    & LAI3g & \autocite{Zhu2013} \\
    & LandFlux-EVAL & \autocite{Mueller2013} \\
    & MLS-Aura & \autocite{Read2015} \\
    & MTE & \autocite{Jung2011} \\
    & NDP & \autocite{Gibbs2006} \\
    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
    \bottomrule
  \end{tabulary}
  \caption{Summary of new recipes and \acs{CMOR}ization scripts for
    observational data contributed to the \acf{ESMValTool} by the author of
    this thesis. The corresponding file names of the recipes in the
    \acs{ESMValTool} repository are given by
    \emph{recipe\_(name\_in\_table).yml}.}
  \label{tab:03:changes_to_esmvaltool}
\end{table}

@@

An important class of recipes in the \ac{ESMValTool} covers international climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which reproduces large parts of the climate model evaluation performed in chapter 9 of the \ac{IPCC}'s \ac{AR5} \autocite{Flato2013}. For this recipe, the two-panel figure 9.42 has been added, which shows the relationship between the historical and pre-industrial \ac{GSAT} and \ac{ECS} (\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS} (\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble. \Cref{fig:03:flato13:a} illustrates that there is no clear connection between \ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high temperatures in the historical or pre-industrial period do no necessarily exhibit a high equilibrium warming in the future or vice versa. On the other hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5 ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this figure only shows a linear regression fit between the two variables and not the expected reciprocal relation.

>>>>

To ensure a rapid evaluation of the \acs{CMIP}6 models, the \ac{ESMValTool} is
fully integrated into the infrastructure of the \ac{ESGF}, which provides the
\ac{CMIP} model output for the general public \autocite{Eyring2016b}. As soon
as new model data is published on the \ac{ESGF} servers, it can be accessed
with the \ac{ESMValTool} and analyzed. This instantaneous evaluation of the
\ac{CMIP} models is urgently needed since there is a growing dependency on
\ac{CMIP} products by a broad research community and by national and
international climate assessments. For this reason, the \ac{ESMValTool} is for
example used to evaluate climate model output and observational data in several
chapters of the upcoming Sixth \acl{AR} (\acs{AR}6) of the \ac{IPCC}.

@@

\begin{table}[!b]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Description \\
    \midrule
    Preprocessor & \emph{amplitude} & Amplitude of periodic phenomena (\eg{}
    cycles) \\
    \cmidrule(l){2-3}
    & \emph{land/sea fraction weighting} & Weighting of fields based on
    the land or sea fraction of the respective grid cells \\
    \cmidrule(l){2-3}
    & \emph{trend} & Linear trend (slope of linear regression) \\
    \midrule
    Derived variable & \emph{asr} & Absorbed shortwave radiation \\
    \cmidrule(l){2-3}
    & \emph{co2s} & Atmospheric \ac{CO2} concentration at
    surface \\
    \cmidrule(l){2-3}
    & \emph{et} & Evapotranspiration \\
    \cmidrule(l){2-3}
    & \emph{rlntcs} & Clear-sky net \acf{TOA} longwave radiation \\
    \cmidrule(l){2-3}
    & \emph{rsntcs} & Clear-sky net \ac{TOA} shortwave radiation \\
    \cmidrule(l){2-3}
    & \emph{uajet} & Position of austral jet stream \\
    \bottomrule
  \end{tabulary}
  \caption{Summary of new preprocessor functions and variable derivation
    scripts contributed to the \acf{ESMValCore} by the author of this thesis.}
  \label{tab:03:changes_to_esmvalcore}
\end{table}

@@

\begin{table}[!b]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Main reference \\
    \midrule
    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
    & \emph{cox18nature} & \autocite{Cox2018} \\
    & \emph{ecs} & \autocite{Gregory2004} \\
    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
    & \emph{tcr} & \autocite{Gregory2008} \\
    \midrule
    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
    & CT2019 & \autocite{Jacobson2020} \\
    & GCP & \autocite{Friedlingstein2019} \\
    & HWSD & \autocite{Wieder2014} \\
    & JMA-TRANSCOM & \autocite{Maki2010} \\
    & LAI3g & \autocite{Zhu2013} \\
    & LandFlux-EVAL & \autocite{Mueller2013} \\
    & MLS-Aura & \autocite{Read2015} \\
    & MTE & \autocite{Jung2011} \\
    & NDP & \autocite{Gibbs2006} \\
    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
    \bottomrule
  \end{tabulary}
  \caption{Summary of new recipes and \acs{CMOR}ization scripts for
    observational data contributed to the \acf{ESMValTool} by the author of
    this thesis. The corresponding file names of the recipes in the
    \acs{ESMValTool} repository are given by
    \emph{recipe\_(name\_in\_table).yml}.}
  \label{tab:03:changes_to_esmvaltool}
\end{table}

@@

An important class of recipes in the \ac{ESMValTool} covers international
climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
reproduces large parts of the climate model evaluation performed in chapter 9
of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
recipe, the two-panel figure 9.42 has been added, which shows the relationship
between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
temperatures in the historical or pre-industrial period do no necessarily
exhibit a high equilibrium warming in the future or vice versa. On the other
hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5
ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
figure only shows a linear regression fit between the two variables and not the
expected reciprocal relation.

### 2021-01-15 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added text and figures to section 4.1

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} is published that includes considerable modifications
compared to the respective predecessor model versions. Currently, we are in a
situation like this with the \acs{CMIP}6 ensemble, for which new data is still
released every day (as of \TheMonth{}). In light of the upcoming international
climate assessment of the \ac{IPCC} (\ac{AR6}), the evaluation of
policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
models is crucial since they provide vital information about the future climate
of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
and analyze climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the
\acs{CMIP}6 models and compare it to corresponding results of predecessor model
generations. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}.

\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}6}
\label{sec:04:evaluation_ecs_and_tcr}

Following \cref{subsec:02:ecs}, \ac{ECS} is calculated with the Gregory
regression method using \ac{GSAT} and \ac{TOA} net radiation data for 150 years
of a \nxcotwo{4} simulation \autocite{Gregory2004}. In contrast to that, the
transient response of the climate system \ac{TCR} is calculated from \ac{GSAT}
data of a \onepctcotwo{} run \commentcite{Bindoff2013}{see
  \cref{subsec:02:tcr}}.

\begin{table}[t]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/CMIP5_ecs-tcr.csv}{}{
    \dataset & \idx & \ecs & \tcr}
  \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
    Details on the calculation on \acs{ECS} and \acs{TCR} are given in
    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively.}
  \label{tab:04:ecs_tcr_cmip5}
\end{table}

\begin{table}[t]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/CMIP6_ecs-tcr.csv}{}{
    \dataset & \idx & \ecs & \tcr}
  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.}
  \label{tab:04:ecs_tcr_cmip6}
\end{table}

>>>>

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} is published that includes considerable modifications
compared to the respective predecessor model versions. Currently, we are in a
situation like this with the \acs{CMIP}6 ensemble, for which new data is still
released every day (as of \TheMonth{}). In light of the upcoming international
climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
models is crucial since they provide vital information about the future climate
of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
and analyze climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the
\acs{CMIP}6 models and compare it to corresponding results of predecessor model
generations. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
author of this thesis contributed their figure 8 (a bar chart showing the
\ac{ECS} for the three climate model generations \acs{CMIP}3, \acs{CMIP}5 and
\acs{CMIP}6), their figure 10 (map plots showing cloud-related feedback
parameters for the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
\cref{fig:04:cloud_feedback_parameters}) and discussion about \ac{ECS} of the
\acs{CMIP}6 models. For \textcite{Meehl2020}, the author of this thesis
contributed all figures and tables of the study and details on the calculations
of \ac{ECS} and \ac{TCR}.

\section{Evaluation of \acs{ECS} and \acs{TCR} over the Years}
\label{sec:04:evaluation_ecs_and_tcr}

Following \cref{subsec:02:ecs}, \ac{ECS} is calculated with the Gregory
regression method using \ac{GSAT} and \ac{TOA} net radiation data for 150 years
of a \nxcotwo{4} simulation \autocite{Gregory2004}. In contrast to that, the
transient response of the climate system \ac{TCR} is calculated from \ac{GSAT}
data at the time of \ac{CO2} doubling in a \onepctcotwo{} run
\commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using these definitions,
we assessed \ac{ECS} and \ac{TCR} for all \acs{CMIP}5 and \acs{CMIP}6 models
where the necessary data is available (as of \TheMonth{}) using the
\ac{ESMValTool}. \Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an
overview over the results for the \acs{CMIP}5 and \acs{CMIP}6 models,
respectively.

\begin{table}[!t]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/CMIP5_ecs-tcr.csv}{}{
    \dataset & \idx & \ecs & \tcr}
  \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
    Details on the calculation on \acs{ECS} and \acs{TCR} are given in
    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively.}
  \label{tab:04:ecs_tcr_cmip5}
\end{table}

\begin{table}[p]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/CMIP6_ecs-tcr.csv}{}{
    \dataset & \idx & \ecs & \tcr}
  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.}
  \label{tab:04:ecs_tcr_cmip6}
\end{table}

The first striking feature of these two tables is the increased \ac{MMM} of
\ac{ECS} and \ac{TCR} in \acs{CMIP}6. For \ac{ECS}, the \acs{CMIP}6 \ac{MMM} is
about $16 \unit{\%}$ ($0.51 \unit{K}$) higher than the corresponding
\acs{CMIP}5 \ac{MMM}. For \ac{TCR}, the relative difference between the two
model ensembles is notably smaller with about $10 \unit{\%}$ ($0.19 \unit{K}$).
The spread in the multi-model ensembles (expressed as the multi-model standard
deviation) shows an even larger increase in the \acs{CMIP}6 ensemble: For
\ac{ECS}, the relative difference is about $49 \unit{\%}$ ($0.72 \unit{K}$ in
\acs{CMIP}5 to $1.07 \unit{K}$ in \acs{CMIP}6) and for \ac{TCR}, the relative
difference is about $17 \unit{\%}$ ($0.36 \unit{K}$ in \acs{CMIP}5 to $0.42
\unit{K}$ in \acs{CMIP}6). The main reason for the increased \ac{MMM} and
spread in the \acs{CMIP}6 ensemble is the existence of several models with very
high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exists a
number of models with very low values of \ac{ECS}. Therefore, the \ac{CMIP}6
model ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding
assessed ranges given by the latest published \acl{AR} of the \ac{IPCC} from
2013 \autocite{Stocker2013} with $\range{1.8}{5.6} \unit{K}$ (\acs{AR}5:
$\range{1.5}{4.5} \unit{K}$) and  $\range{1.3}{3.0} \unit{K}$ (\acs{AR}5:
$\range{1.0}{2.5} \unit{K}$) for \ac{ECS} and \ac{TCR}, respectively.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
    and the subsequent \acp{AR} of the \acf{IPCC}. Orange and green bars
    represent the modeled ranges of \acs{ECS} and \acs{TCR}, respectively, from
    the different climate model ensembles of the \acf{CMIP} with the
    corresponding \acf{MMM} illustrated with the horizontal black lines (the
    green dot represents output from a single climate model). The numbers
    correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
    \textcite{Meehl2020}.}
  \label{fig:04:historical_ecs_tcr_ranges}
\end{figure}

To illustrate this and put it into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
\unit{K}$ has almost remained unchanged during the last 40 years (with the
exception of \acs{AR}4, where the lower bound was temporarily increased to $2.0
\unit{K}$). For \acs{AR}2, \acs{AR}4 and \acs{AR}5, the corresponding climate
model generations from \acs{CMIP}1, \acs{CMIP}5 and \acs{CMIP}6 more or less
agree with this assessed range. In contrast to that, some of the early climate
models used in \acs{AR}1 and the \acs{CMIP}2 models used in \acs{AR}3 exhibit
values of \acs{ECS} above $4.5 \unit{K}$, resulting in a upper model range of
about $0.6 \unit{K}$ higher than the assessed range. However, these deviations
from the assessed range are negligible to the ones present in the \acs{CMIP}6
ensemble, which have become evident over the course of 2020. As shown in
\cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper bounds are
more extreme than in any previous climate model generation, resulting in a
\acs{CMIP}6 model range of $\range{1.8}{5.6} \unit{K}$ (as of \TheMonth{}).
This increase in range is particularly relevant for the upper bound, which is
about $1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
Moreover, not only a small fraction of models exceeds this upper bounds: out of
the 42 analyzed \ac{CMIP}6 models, a third (14 models) has an \ac{ECS} above
$4.5 \unit{K}$. Because of the massive political and societal relevance of
\ac{ECS}, its apparent increase in the \ac{CMIP}6 ensemble is currently one of
the most important questions for the entire climate modeling community. A list
of possible reasons for this increased climate sensitivity is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

\section{Possible Reasons for High Climate Sensitivity in \acs{CMIP}6}
\label{sec:04:possible_reasons_high_ecs_cmip6}
